---
title: "General Additive Models"
author: "Manisha"
date: "2023-11-09"
output: html_document
---


## load data and main libraries
```{r}
setwd("C:\\Users\\manis\\Documents\\Fall3\\ML\\Notes")
cement = read.csv("cement.csv")
library(dplyr)
library(ggplot2)
View(cement)
summary(cement)
```

##Piecewise linear regression
## predicting STRENGTH using the RATIO variable as well the X2STAR variable which is (x1-k)x2
 
```{r cars}
cement.lm <- lm(STRENGTH ~ RATIO + X2STAR, data = cement)

summary(cement.lm)
```

## Including Plots
## in the plot below how above the knot value of 70, the slope and intercept of the regression line changes
```{r plot, echo=FALSE}
ggplot(cement, aes(x = RATIO, y = STRENGTH)) +
  geom_point() +
  geom_line(data = cement, aes(x = RATIO, y = cement.lm$fitted.values)) +
  ylim(0,6)
```
##previous example dealt with piecewise functions that are continuous
##model to make the linear discontinuous:
## with addition of x2 line is no onger attached
```{r}
cement.lm <- lm(STRENGTH ~ RATIO + X2STAR + X2, data = cement)

summary(cement.lm)

qplot(RATIO, STRENGTH, group = X2, geom = c('point', 'smooth'), method = 'lm', data = cement, ylim = c(0,6))
```
###########MARS(EARTH)######################################

#Garage no straight forward relation with Saleprice ( from Ames housing data)
```{r}
ggplot(training, aes(x = Garage_Area, y = Sale_Price)) +
  geom_point()
```

##earth package is the earth function.
```{r}
library(earth)

mars1 <- earth(Sale_Price ~ Garage_Area, data = training)
summary(mars1)
```
##In above output 6 pieces defined by 5 knots ( coefficients attached to each of those pieces are the same as what we would have in piecewise linear regression) , at the bottom we have Generalized Rsquare and Rsquare value

--

## visualising piecewise relation with predicted values

```{r}
ggplot(training, aes(x = Garage_Area, y = Sale_Price)) +
  geom_point() +
  geom_line(data = training, aes(x = Garage_Area, y = mars1$fitted.values), color = "blue")
```
##building mars/earth on all variables in the data set
```{r}
mars2 <- earth(Sale_Price ~ ., data = training)
summary(mars2)
```
## In the above output knot values are defined for all the varaibles that are in the model
## 10 out of 14 orginal variables are used in final model
## we have variable importance in the bottom, Rsquare & GRsquare value increased compared to single variable model, 
## as other variables are added, Garage Area has less knots now, algorithm prunes the knots

##variable importance
##MARTH ranks variables by how many of these “best models” of each size that variable appears in. The more subsets (or “best models”) that a variable appears in, the more important the variable. 
```{r}
evimp(mars1)
```
##nsubsets - number of sets variable appears in
----------

###gam function from mgcv package and with Spline function on predictors
```{r}
library(mgcv)

gam1 <- mgcv::gam(Sale_Price ~ s(Garage_Area), data = training)
summary(gam1)
#a section for coefficients that are not involved in splines and a section for smoothing terms.
#p-value attached to the spline of Garage_Area shows the significance of that variable to the model as a whole.
```

##plot function (to see 'relationship between the variable and its predictions of the target. ')
```{r}
plot(gam1)
```

## GAM on all varaibles
## categorical variables are entered as either character variables or with the factor function. The continuous variables are defined with the spline function s
```{r}
gam2 <- mgcv::gam(Sale_Price ~ s(Bedroom_AbvGr, k = 5) + 
                               s(Year_Built) +
                               s(Mo_Sold) +
                               s(Lot_Area) +
                               s(First_Flr_SF) + 
                               s(Second_Flr_SF) +
                               s(Garage_Area) +
                               s(Gr_Liv_Area) +
                               s(TotRms_AbvGrd) +
                               Street +
                               Central_Air +
                               factor(Fireplaces) + 
                               factor(Full_Bath) +
                               factor(Half_Bath)
                  , method = 'REML', data = training)
summary(gam2)
#top half of the output has the variables not in splines, while the bottom half has the spline variables.
#**some variables with high p-values that could be removed from the model. 
##select = TRUE then the model will penalize variables’ edf values. You can think of an edf value almost like a polynomial term.
##The selection technique will zero out this edf value - essentially, zeroing out the variable itself.

## here 'Mo_Sold variable is essentially zeroed from the model.'
```




